Type Flattening Obfuscation

### 2020-01-24 paper/tool.tex
implement type equivalence

\usepackage{cleveref}

@@

\emph{First}, type obfuscation can be an indirect effect of data or control flow obfuscation.
Indeed, type reconstruction algorithms need both data and control flow to build type constraints,
if any of them is hidden
then the algorithms will not work. Or if the function boundary cannot be found
(because of anti-disassembing tricks, for example), then there are no inputs (i.e. high-level
objects) for type reconstruction. \emph{Second}, high-level types seem too coarse to be worthy
of being protected, in many cases just knowing certain values of the input which make the program
exploitable is enough. But knowledge about types enlarges attack surfaces because more analysis
can be proceeded, beside decompilation see the survey~\cite{caballero_type_2016} for a
more complete list.
% We exploit the \emph{semantics gap}

\section{Type flattening compiler}

We introduce \emph{ucc}, a proof-of-concept C compiler that explicitly obfuscates high-level types.
We attacks on the core of type reconstruction algorithms, that is the data flow used to build
type constraints, and the \emph{semantic gap} between types in the high-level language and their
representations in low-level machine code.

The compiler is open source and implemented in Rust. The source code, an user guide as well as some
demo results are available at (cite).

>>>>

% \usepackage{cleveref}
% \usepackage{adjustbox}
\usepackage[strict]{changepage}

@@

\emph{First}, type obfuscation can be obtained as side effects of data or control flow obfuscation.
Indeed, type reconstruction algorithms need both data and control flow to build type constraints,
if any of them is hidden
then the algorithms cannot work correctly. Or if the function boundary cannot be found
(because of anti-disassembing tricks, for example), then there are no inputs (i.e. high-level
objects) for type reconstruction. \emph{Second}, high-level types seem too coarse to be worthy
of being protected, in many cases just knowing certain values of the input which make the program
exploitable is enough. But knowledge about types expands attack surfaces because more analysis
can be proceeded, beside decompilation see the survey~\cite{caballero_type_2016} for a
more complete list.
% We exploit the \emph{semantics gap}

\section{Type flattening compiler}
We present \emph{uCc}, an open source C compiler that explicitly obfuscates high-level types.
Our effort is to attack the core of type reconstruction algorithms,
that is the data flow used to build type constraints, and the \emph{semantic gap} between types in the high-level language and their
representations in low-level machine code. For that purpose, \emph{uCc} uses extensively
obfuscation transformations based on Mixed Boolean Arithmetic expressions~\cite{eyrolles_obfuscation_2017,zhou_information_2007}.
To increase code diversity, it is a \emph{probabilistic} compiler: given the source
code, \emph{uCc} generates each time a different (but computationally equivalent) output machine code.

The compiler is implemented in Rust, uses \emph{cranelift}~\cite{noauthor_cranelift_nodate} as
backend. The source code and a short user guide are at (cite). We test outputs of
\emph{uCc} against decent decompilers, table~\ref{flattening} show experimental results

\begin{table}
	\begin{center}
		\caption{Functions and original types}
		\begin{tabular}{|l|l|}
			\hline
			Description & Function type \\
			\hline
			identity & \texttt{int id(int)} \\
			\hline
			division & \texttt{int div(short, char)} \\
			\hline
			modular & \texttt{int mod(int)} \\
			\hline
			increase pointer then dereference & \texttt{int inc\_deref(int*)} \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

\begin{table}
	\begin{adjustwidth}{-6.21in}{-5.95in}
		\begin{center}
			% \centering
			\caption{Type flattening effect}\label{flattening}
			% \begin{adjustbox}{max width=1.5\textwidth}
			\begin{tabular}{|l|l|l|l|l|}
				\hline
				Original type & Hex-Rays & JEB & Ghidra & Snowman \\
				\hline
				\texttt{int id(int)} & \texttt{int64 id(int64)} & \texttt{ulong id(ulong)} &
				\texttt{ulong id(void)} & \texttt{int64 id(int64)} \\
				\hline
				\texttt{short div(short, char)} & \texttt{int64 div(uint64, uint64)} & \texttt{div\_t div(int, int)} &
				\texttt{div\_t div(int, int)} & \texttt{int64\_t div(int64, int64)} \\
				\hline
				\texttt{char mod(short, char)} & \texttt{int64 mod(int64, int64)} & \texttt{ulong mod(ulong, ulong)} &
				\texttt{ulong mod(void)} & \texttt{int64 mod(int64, int64)} \\
				\hline
				\texttt{int inc\_deref(int*)} & \texttt{int64 inc\_deref(int64)} & \texttt{ulong inc\_deref(ulong)} &
				\texttt{ulong inc\_deref(undefined8)} & \texttt{int64 inc\_deref(uint64)} \\
				\hline
			\end{tabular}
		\end{center}
	\end{adjustwidth}
\end{table}
